Optimal (Multiway) Spatial Joins

### 2024-06-01 msj.tex
yufei

Given an $(n_1, n_2, ..., n_k)$-input $\set{R_1, R_2, ..., R_k}$, we execute the algorithm by traversing a root-to-leaf path in $\T$. In general, at an internal node with the instruction $(i, \lambda, \side; i', \lambda', \side')$, we first identify
\myitems{
    \item $r$ as the rectangle with ID $\lambda$ in $R_i$;
    \item $r'$ as the rectangle with ID $\lambda'$ in $R_{i'}$.
}
Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
\myitems{
    \item the left child if $\side(r) < \side'(r')$;
    \item the middle child if $\side(r) = \side'(r')$;
    \item the right child if $\side(r) > \side'(r')$.
}
At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in that result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$. The algorithm is {\em correct} if it produces the correct output on all $(n_1, n_2, ..., n_k)$-inputs. Its {\em worst-case cost} is the height of $\T$.

@@

Through a reduction from the element distinctness problem \cite{dl79}, one can show that, for $k = 2$, any comparison-based algorithm capable of solving $(n/2, n/2)$-inputs must have a worst-case cost of $\Omega(n \log n)$.

@@

\noindent \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$. Let $h$ be the full segment of $h'$. Following our conventions, $B_\bm{t}$ denotes the intersection point of $h'$ and $v$, and it is covered by the rectangles $r_1,...,r_{k-3}$. As $h$ is the full segment of $h'$, we have $B_\bm{t} \in h$, and $B_\bm{t}$ is also the intersection of $h$ and $v$. Let $r_{k-2}$ be an arbitrary rectangle in $\contained_{R_{k-2}}(\bm{t})$. By the definition of $\contained_{R_{k-2}}(\bm{t})$, $r_{k-2}$ covers the effective horizontal segment of $h'$, whose right enpoint is $B_{\bm{t}}$. Therefore, $B_{\bm{t}}$ is covered by each of the rectangles $r_1,...,r_{k-2}$. Equivalently, $h \cap v \bigcap_{i = 1}^{k-2} r_i = B_\bm{t} \neq \emptyset$. Hence,
$\bm{t} \in \J(R_1,...,R_{k-2},H,V)$. To show that $\bm{t} \in \J_2$, we only need to prove that $r_{k-2}$ covers the left endpoint of $h$. It is true because $r_{k-2}$ covers the effective horizontal segment of $h'$, whose left endpoint is the same as the left endpoint of $h$.

@@

\noindent \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, $\bm{t}$ be as defined in the statement. First, we prove that $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$. It suffices to show that the intersection of $h'$ and $v$ is contained in $r_1,...,r_{k-3}$. We will prove this by showing that $h'\cap v$ is the same as $h\cap v$, which is covered by $r_1,...,r_{k-2}$, because $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.

>>>>

Given an $(n_1, n_2, ..., n_k)$-input $\set{R_1, R_2, ..., R_k}$, we execute the algorithm by traversing a root-to-leaf path in $\T$. In general, at an internal node with the instruction $(i, \lambda, \side; i', \lambda', \side')$, we first identify
\myitems{
    \item $r$ as the rectangle with ID $\lambda$ in $R_i$;
    \item $r'$ as the rectangle with ID $\lambda'$ in $R_{i'}$.
}
Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
\myitems{
    \item the left child if $\side(r) < \side'(r')$;
    \item the middle child if $\side(r) = \side'(r')$;
    \item the right child if $\side(r) > \side'(r')$.
}
At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in that result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$. The algorithm must produce the correct output on all $(n_1, n_2, ..., n_k)$-inputs. Its {\em worst-case cost} is the height of $\T$.

@@

Through a reduction from element distinctness \cite{dl79}, one can show that, for $k = 2$, any comparison-based algorithm correctly solving all $(n/2, n/2)$-inputs must have a worst-case cost of $\Omega(n \log n)$.

@@

\extraspacing \underline{\em Proof of Statement (1).} Consider any tuple $\bm{t} = (r_1,...,r_{k-3}, h', v) \in \J(R_1,...,R_{k-3},H',V)$. Let $h$ be the full segment of $h'$. The rectangle $B_\bm{t}$ defined in \eqref{eqn:B_t} is the intersection point of $h'$ and $v$ (this point is covered by all of $r_1,...,r_{k-3}$). As $h$ is the full segment of $h'$, the point $B_\bm{t}$ must also be the intersection point of $h$ and $v$.

@@

Let $r_{k-2}$ be an arbitrary rectangle in the set $\contained_{R_{k-2}}(\bm{t})$. The definition of $\contained_{R_{k-2}}(\bm{t})$ tells us that $r_{k-2}$ covers the effective horizontal segment of $h'$ (see Section~\ref{sec:hv:type2} for the definition of ``effective horizontal segment''), whose right endpoint is $B_{\bm{t}}$. This means that $B_\bm{t}$ falls in $r_{k-2}$ as well.

Now we know that $h \cap v \cap \bigcap_{i = 1}^{k-2} r_i$ contains $B_\bm{t}$ and is thus non-empty. Hence,
$(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J(R_1,...,R_{k-2},H,V)$. To show $(r_1,...,r_{k-3}, r_{k-2}, h, v) \in \J_2$, we still need to prove that $r_{k-2}$ covers the left endpoint of $h$. This follows immediately from the fact that $r_{k-2}$ covers the effective horizontal segment of $h'$ (which shares the same left endpoint as $h$).

\extraspacing \underline{\em Proof of Statement (2).} Let $(r_1,...,r_{k-2},h,v)$, $h'$, $\bm{t}$ be as defined in the statement. First, we prove that $\bm{t}\in \J(R_1,...,R_{k-3}, H', V)$. It suffices to show that the intersection of $h'$ and $v$ is contained in $r_1,...,r_{k-3}$. We will prove this by showing that $h'\cap v$ is the same as $h\cap v$, which is covered by $r_1,...,r_{k-2}$, because $\bm{t} \in \J(R_1,...,R_{k-2},H,V)$.
